---
title: "Comparing Expected Lifetime of LED Bulbs under Temperature S and Temperature T"
subtitle: "STA 305 Assignment 2"
author: "Yu-Chun Chien"
date: "2/14/2021"
output: pdf_document
---

# Part I

## Treatment S
Generate 9 observations from exp($\lambda = \frac{1}{3}$) to correspond to treatment S.

```{r}
set.seed(1005194380)
rexp(9, rate=1/3)
```
Observed values in the order they appeared: 1.340, 4.024, 1.436, 3.568, 8.344, 3.833, 2.975, 3.406, 2.715

## Treatment T
Generate 9 observations from exp($\lambda = 1$) to correspond to treatment T.

```{r}
set.seed(1005194380)
rexp(9, rate=1)
```

Observed values in the order they appeared: 0.447, 1.341, 0.479, 1.189, 2.781, 1.278, 0.992, 1.135, 0.905

## Pairs of Observations of Treatment S and T

```{r}
treat_s_yc <- c(1.340, 4.024, 1.436, 3.568, 8.344, 3.833, 2.975, 3.406, 2.715)
treat_t_yc <- c(0.447, 1.341, 0.479, 1.189, 2.781, 1.278, 0.992, 1.135, 0.905)
pair_no_yc <- c(1, 2, 3, 4, 5, 6, 7, 8, 9)
random_pair_yc <- data.frame(pair_no_yc, treat_s_yc, treat_t_yc)
random_pair_yc
```

# Part II

## Completely Randomized Design

### Randomization Distribution 

```{r, warning=F}
# pool data
pair_yc <- c(treat_s_yc, treat_t_yc)
N_yc <- choose(18, 9)
diff_yc <- numeric(N_yc)

library(combinat)
index_yc <- combn(1:18,9)

for (i in 1:N_yc){
  diff_yc[i] <- mean(pair_yc[index_yc[,i]]) - mean(pair_yc[-index_yc[,i]])
}
```

Number of Values this Distribution Contains: 48620

Probability of the Observed Treatment Allocation: $\frac{1}{48620}$

### Histogram of the Randomized Distribution

```{r}
hist(diff_yc, xlab= "|meanS-meanT|", main = "Randomization Distribution CRD 4380")

observe_yc <- mean(treat_s_yc) - mean(treat_t_yc)

abline(v=observe_yc, col="blue")
abline(v=-(observe_yc), col="blue")
text(2.2, 8600, "2.34", col = "blue")
text(-2.1, 8600, "-2.34", col = "blue")
```

### Randomization Test

$H_0: \mu_s = \mu_t$
$H_A: \mu_s \neq \mu_t$

```{r}
# p-value
tbar_yc <- mean(diff_yc)

p_yc <- sum(abs(diff_yc - tbar_yc)>=abs(observe_yc - tbar_yc))/N_yc
round(p_yc, 2)
```

The p-value is 0. According to the randomization test, assume there is no difference in life expectency of the bulbs between temperature S and temperature T, approximately 0% of randomizations would produce an extreme or more extreme difference than the observed mean differenence $|2.34|$ (probability = the area to the right of 2.34 and to the left of -2.34 in the histogram). Thus, the result of the randomization test indicates that it is quite unlikely that light bulbs will have the same life expectancy under temperature S and T.

## Randomized Paired Design

### Randomization Distribution

Number of Values this distribution contains: $2^9=512$

Probability of the observed treatment allocation: $\frac{1}{512}$

### Histogram of Randomization Distribution


```{r}
diffp_yc <- treat_s_yc - treat_t_yc
mean_diffp_yc <- mean(diffp_yc)

Np_yc <- 2^(9)
nump_yc <- numeric(Np_yc)
mult_yc <- list(c(-1,1))
assign_yc <- expand.grid(rep(mult_yc, 9))

for (i in 1 : Np_yc){
  nump_yc[i] <- mean(as.numeric(assign_yc[i,] * diffp_yc))
}
```

```{r}
# histogram
hist(nump_yc, xlab = "Mean Difference", main = "Randomization Distribution Randomized Paired 4080")
abline(v = mean_diffp_yc, col="blue")
abline(v = -(mean_diffp_yc), col="blue")
text(2.15, 97, "2.34", col = "blue")
text(-2.1, 97, "-2.34", col = "blue")
```

### Randomization Test

$H_0: \mu_s = \mu_t$
$H_A: \mu_s \neq \mu_t$

```{r}
# p-value
tbarp_yc <- mean(nump_yc)

p_val_yc <- sum(abs(nump_yc - tbarp_yc)>=abs(mean_diffp_yc - tbarp_yc))/Np_yc
round(p_val_yc, 2)
```

The p-value is 0. According to the randomization test, assume there is no difference in life expectency of the bulbs between temperature S and temperature T, approximately 0% of randomizations would produce an extreme or more extreme difference than the observed mean differenence $|2.34|$ (probability = the area to the right of 2.34 and to the left of -2.34 in the histogram). The result of the randomization test indicates that it is quite unlikely that light bulbs will have the same life expectancy under temperature S and T.

# Part III

## Two-Sample T-test for Completely Randomized Design

```{r}
t.test(treat_s_yc, treat_t_yc, var.equal = F, alternative = "two.sided")
```
p-value of completely randomized design: 0

According to the two-sample t-test, the probability that the mean of the two measurements is equal is 0. Using a benchmark significance level of 5%, it is quite unlikely that the mean of the two measurements is equal. Instead, the mean of the expected lifetime under temperature S is larger than the mean of the expected lifetime under temperature T.

## Paired T-test for Randomized Paired Design

```{r}
t.test(treat_s_yc, treat_t_yc, paired = T, alternative = "two.sided")
```
p-value of randomized paired design: 0

According to the paired t-test, the probability that the mean of the paired differences is zero is approximately 0. Using a benchmark significance level of 5%, it is quite unlikely that the mean of the paired difference is zero. Instead, the mean of the difference is greater than zero; on average, the expected lifetime under temperature S is larger than the expected lifetime under temperature T. 

## Assumption behind T-test

The assumption of independence is satisfied while the assumption of normality is not satisfied 

(1) The two samples (treatment S and T) are sampled individually without influencing each other, which makes them independent of each other.

(2) The two samples are not normally distributed. According to the Normal QQ plots, the two samples are right-skewed and do not follow a normal distribution.
```{r}
# normality
par(mfrow=c(1,2))
qqnorm(treat_s_yc, main = "Normal Q-Q Plot Treatment S")
qqline(treat_s_yc)
qqnorm(treat_t_yc, main = "Normal Q-Q Plot Treatment T")
qqline(treat_t_yc)
```

### Result of T-test vs. Randomization Test

Yes, the results of the t-test agree with the results of the randomization test. Both tests have p-value that is equal to zero, which suggested that there is a difference in the light bulb life expectancy under temperature S and temperature T. 

# Part IV

## Power of Completely Randomized Design and T-test

```{r}
set.seed(1005194380)

pvals_yc <- replicate(1000, t.test(rexp(20, 1/3), rexp(20, 1), var.equal = F, alternative = "two.sided")$p.value)

sum(pvals_yc <= 0.05)/1000

```

Power: 0.86

## Randomized Paired Design and T-test

```{r}
set.seed(1005194380)

ppair_yc <- replicate(1000, t.test(rexp(20, 1/3), rexp(20, 1), paired = T, alternative = "two.sided")$p.value)

sum(ppair_yc <= 0.05)/1000
```

Power: 0.85

## Completely Randomized Design and Wilcoxon Test

```{r}
set.seed(1005194380)

wil_crd_yc <- replicate(1000, wilcox.test(rexp(20, 1/3), rexp(20, 1), alternative = "two.sided")$p.value)

sum(wil_crd_yc <= 0.05)/1000
```

Power: 0.80

## Randomized Paired Design and Wilcoxon Test

```{r}
set.seed(1005194380)

wil_paired_yc <- replicate(1000, wilcox.test(rexp(20, 1/3), rexp(20, 1), paired = T, alternative = "two.sided")$p.value)

sum(wil_paired_yc <= 0.05)/1000
```

Power: 0.83

## Comparison of Power 

Power that is larger than 0.80 is desirable. Here, the comparison of power is as follows:

Completely Randomized Design and T-test (0.86) > Randomized Paired Design and T-test (0.85) > Randomized Paired Design and Wilcoxon Test (0.83) > Completely Randomized Design and Wilcoxon Test (0.80)

## Checking for Normality

```{r}
set.seed(1005194380)

# treatment S
qqnorm(rexp(20, 1/3), main = "Normal Q-Q Plot Treatment S")
qqline(rexp(20, 1/3))

# treatment T
qqnorm(rexp(20, 1), main = "Normal Q-Q Plot Treatment T")
qqline(rexp(20, 1))

# treatment S + T
combine_yc <- c(rexp(20, 1/3), rexp(20, 1))
qqnorm(combine_yc, main = "Normal Q-Q Plot Treatment S + T")
qqline(combine_yc)
```

### Statistical Test for Completely Randomized Design

For the completely randomized design, we need both samples (treatment S and treatment T) to follow a normal distribution for carrying out the two-sample t-test. However, the normal QQ plot suggested that although the 20 samples of treatment S follow a normal distribution, the 20 samples of treatment T does not follow a normal distribution. Thus, the Mann-Whitney Test would be more suitable for the completely randomized design.

### Statistical Test for the Randomized Paired Design

For the randomized paired design, we need the combined dataset (20 samples of treatment S and 20 samples of treatment T) to follow a normal distribution for carrying out the paired t-test. However, according to the normal QQ plot, the combined dataset does not follow a normal distribution. Therfore, paired t-test is not applicable here and we should instead conduct a Mann-Whitney Test.
